---
title: "Prediction Assignment Writeup"
output: document
---
#loading and processing
```{r warning=FALSE,cache=TRUE}
setwd("C:/Users/PurushothamNaidu/Documents")
url_train="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url_test="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(url_train,destfile="training_data.csv")
download.file(url_test,destfile="testing_data.csv")
training<-read.csv("C:/Users/PurushothamNaidu/Documents/training_data.csv",na.strings = c("NA", "#DIV/0!", ""))
testing <- read.csv("C:/Users/PurushothamNaidu/Documents/R/testing_data.csv",na.strings = c("NA", "#DIV/0!", ""))
```

#Removing 
```{r warning=FALSE,cache=TRUE}
training <- training[, which(colSums(is.na(training)) == 0)] 
testing <- testing[, which(colSums(is.na(testing)) == 0)]
training <- training[,-c(1:7)] ##the first 7 columns are variables that has no relationship with "class"
testing <- testing[,-c(1:7)]
``
#Partioning the training set into training and crossvalidation datasets
```{r warning=FALSE, cache=TRUE}
library(caret)
set.seed(888)
training = data.frame(training)
inTrain <- createDataPartition(training$classe, p=0.70, list=F)
train <- training[inTrain, ]
validation <- training[-inTrain, ]
```
#Modelling with regression tree ("rpart")
```{r warning=FALSE, cache=TRUE}
fit1 <- train(classe ~ ., method="rpart", data=train)
val1 <- predict(fit1, validation)
confusionMatrix(validation$classe, val1)
```
#Modelling with random forest ("rf")
```{r warning=FALSE, cache=TRUE}
fit2 <- train(classe ~ ., method="rf", data=train, prox=TRUE,ntree=250)
val2 <- predict(fit2, validation)
confusionMatrix(validation$classe, val2)
```

#Modelling with boosted trees ("gbm")
```{r warning=FALSE, cache=TRUE}
fit3 <- train(classe ~ ., method="gbm", data=train,trControl=trainControl(method = "repeatedcv", number = 5, repeats = 1),verbose=FALSE)
val3 <- predict(fit3, validation)
confusionMatrix(validation$classe, val3)
```
#Prediction
We used the random forest model for prediction
```{r warning=FALSE, cache=TRUE}
pred <- predict(fit2, newdata=testing)
pred
```
#Plotting decision tree(method="rpart")
```{r warning=FALSE, echo=FALSE,cache=TRUE}
plot(fit1$finalModel, uniform=TRUE,main="Classification Tree")
text(fit1$finalModel, use.n=TRUE, all=TRUE, cex=0.8)
```
